---
title: "FlowerPower"
author: "Sehaj Raina"
date: "March 7, 2019"
output: html_document
---

<br><br>

1. Load packages
```{r}
suppressMessages(library(vegan))
suppressMessages(library(ape))
suppressMessages(library(ggplot2))
suppressMessages(library(ggtree))
```

<br>

2. Importing and data wrangling 
```{r}
samples<-read.csv("flower.csv")

#change population from continuous to a categorical variable 
str(samples$Population)
samples$Population<-as.factor(samples$Population)

samples1<-read.csv("flower.csv", header=T, row.names="Quadrate")

table1<- samples1[, -c(1:5)] #do not need these columns for analysis 

x<-rowSums(table1)
sum(x)
Drop<-rowSums(table1) < 1
sum(Drop) #nothing to drop
```

<br>

3. Visualization by Location - Binary Distance Matrix
```{r}
table1_bin<-table1 #convert to binary
table1_bin[table1_bin>0]<-1

table1_bin_dist<-dist(table1_bin, method="binary")

table1_tree<-nj(table1_bin_dist)
ggtree(table1_tree, layout="rectangular") %<+% samples + 
  geom_tiplab(aes(colour=Location)) + 
  theme(legend.position="right")
```

### __Figure 1.__ *Binary Distance Matrix Tree by Location*

<br>

4. Visualization by Location - Dissimilarity
```{r}
table1_dist<- vegdist(table1) #use the data WE DIDNT transform to binary

table1_tree2<-nj(table1_dist)
ggtree(table1_tree2,layout="rectangular") %<+% samples + 
  geom_tiplab(aes(color=Location)) + 
  theme(legend.position="right")
```

### __Figure 2.__ *Dissimilarity Tree by Location*

<br>

5. Visualization by Location - NMDS
```{r}
NMDSdat<-metaMDS(table1_dist, k=2)

PDat<-data.frame(NMDS1=NMDSdat$point[,1],
                 NMDS2=NMDSdat$point[,2],
                 Quadrate=row.names(table1))
 
PDat<-merge(PDat,samples,by="Quadrate",all.x=T,all.y=F) #join data sets 

qplot(x=NMDS1,y=NMDS2,colour=Location,alpha=I(0.8),data=PDat)+theme_bw()
```

### __Figure 3.__ *Non-Metric Multidimensional Scaling (NMDS) by Location*

<br>

6. Visualization by Population - Binary Distance Matrix
```{r}
table3_tree<-nj(table1_bin_dist)
ggtree(table3_tree, layout="rectangular") %<+% samples + 
  geom_tiplab(aes(colour=Population)) + 
  theme(legend.position="right") 
```

### __Figure 4.__ *Binary Distance Matrix Tree by Population*

<br>

7. Visualization by Population - Dissimilarity
```{r}
table4_tree2<-nj(table1_dist)
ggtree(table1_tree2,layout="rectangular") %<+% samples + 
  geom_tiplab(aes(color=Population)) + 
  theme(legend.position="right")
```

### __Figure 5.__ *Dissimilarity Tree by Population*

<br>

8. Visualization by Population - NMDS
```{r}
qplot(x=NMDS1,y=NMDS2,colour=Population,alpha=I(0.8),data=PDat)+theme_bw()
```

### __Figure 6.__ *Non-Metric Multidimensional Scaling (NMDS) by Population*

<br>

9. Visualization by Location and Population - NMDS
```{r}
qplot(x=NMDS1,y=NMDS2,shape=Population, colour=Location, alpha=I(0.8),data=PDat)+theme_bw()
```

### __Figure 7.__ *Non-Metric Multidimensional Scaling (NMDS) by Location and Population*

<br>

####




